---
title: "WUP_04"
author: "Andrew Valles Tovar"
date: "Spring 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file = "GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(mydata,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot()
```

Just by looking, the high school and LT high school box participants would perform lower than the junior college and graduate participants. When the high school box and LT high school boxes are moved to a different category, their placements will not merge with the boxes placed in junior college or graduate boxes, indicating that the results are unique.
  
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```
The Q1 in both LT high school and high school categories are both the same. Meanwhile, the Q3 of the graduate category is higher than the Q1 of LT high school and high school categories.
  
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```
The probability of this data having no difference is 0.000236%. The probability shows that there is a difference, but a multiple comparison test is needed to truly find out the differences.
  
```{r}
a <- TukeyHSD(model1, conf.level=.95)
par(mar = c(3, 15.7, 2, 2))
plot (a, las = 2)
```

What the graph shows here is that when running the TukeyHSD test, both the high school and graduate high school categories have statistical differences at 95% and above confidence level.
  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
